\textbf{\sc Towards Deliberative Control in Marine Robotics}

### 2012-02-22 trunk/concepts.tex
added figures missing in concepts.tex

  \item \textit{state}: Tokens can be \texttt{ACTIVE, INACTIVE,
      MERGED}, or \texttt{REJECTED}. The state variable captures the
    token's current state and its reachable states through further
    restriction.  This is required to support \texttt{CAIP}'s approach
    to planning where goals and subgoals can be satisfied by either
    adding new intervals (tokens) to the plan, or by merging with
    existing intervals. % Token state lifecycle is examined in detailed
    % in a separate section below.

@@

\item[\textbf{Built-in Object Types}] In a domain model, there may be
  some classes that don't have any time-dependent state or behavior,
  and therefore those classes will not have any token types associated
  with them. For instance in the Shopping Agent model, the set
  locations and products are static for the agent's purposes so one
  may only want to declare them but not associate any state or
  behavior with them.  However, a common case in a non-trivial domain
  model is that objects will have many associated tokens in order to
  describe their state and behavior throughout a plan. There are some
  object types that are commonly used in domain descriptions, that \eu
  provides a built-in implementation for them:

@@

\item[\textbf{Token State Model}] In \eus representation, goals are
  temporally scoped states that need to be achieved, or actions that
  must be performed. As a result, \eu internally represents goals as
  tokens.  Goals can be posted explicitly, for example, when a user
  asks the Shopping Agent to own a specific product by a specific
  time, or implicitly when subgoals are created as a result of domain
  compatibilities \kcomment{Define a compatibility}. For instance,
  when the Shopping Agent must get to a location where a product is
  available before it is able to purchase it: in this case the
  original goal of purchasing the product spawns a subgoal of being at
  a specific location without which the purchase will not occur as
  defined in the domain model.  Looking at a how a simple Shopping
  Agent goal would be addresed by \eu helps illustrate how Token
  States support the planning process.  Let's assume the planner
  starts with a partial plan that places the Shopping agent at home at
  time $0$ with no products in possession.  If a goal is posted for
  the Agent to own a drill by time 10. \kcomment{Use Fig 4 and edit::
    Add figure showing an AgentLocation timeline with an At(Home)
    token with start=0 and end open, an empty AgentBag timeline, and a
    Own(Drill) token with start=10 floating below both timelines}

  Note that the \texttt{Own(Drill)} token that represents the goal is
  not immediately placed on the \texttt{AgentBag} timeline since \eu
  must determine whether it can achieve that goal and how. When the
  goal is posted and before any decisions are made, the
  \texttt{Own(Drill)} token is created in an \texttt{INACTIVE}
  state. \eu then considers the partial plan and evaluates two
  possible alternatives:

@@

When a token is merged, all restrictions imposed on the merged token
are transferred to the active token upon which it is merged. Merging a
token requires finding a target active token that is compatible. Two
tokens are merge-compatible if they are instances of the same
predicate and that no intersections between corresponding variables
are empty. The effect of merging is illustrated in Figure 3
\kcomment{Which figure?}.\comment{Use previous fig and modify as
  appropriately:: Add figure showing the effect of merging; all figs
  on this section are conjoined. JB to sketch it out.}

When a token is activated all compatibilities \kcomment{this should
  have been defined earlier} associated with its type are evaluated,
which may lead to new subgoals being generated. In our example, there
could be a compatibility stating that the Shopping Agent must be at
the same location as the product it is purchasing. In such a case,
activating the \texttt{Own(Drill)} token will cause a new
\texttt{At(Drill.location)} to be created in an \texttt{INACTIVE}
state, which in turn could spawn its own subgoals.

\comment{Use previous fig and modify as appropriately: Add figure
  showing the effect of activating the token by putting it on the
  AgentBag timeline, and a new At(Drill.location) token floating
  below}

If neither activating nor merging work, a token may be
\texttt{REJECTED}, this is possible only for goals posted by the end
user and specified as optional. Subgoals generated as a result of
token activation cannot be rejected; if a subgoal cannot be satisfied
through activation or merging then it will cause the original
activation decision that spawned it to be retracted.

As can be seen from the above example, a token's state can go from
\texttt{INACTIVE} to \texttt{ACTIVE, MERGED} or \texttt{REJECTED}, as
shown in figure below \kcomment{Which figure?}.

\comment{Maybe: Add figure showing possible state transitions. Need a FSM.}

@@

  \comment{Use previous fig and modify as appropriately (used as
    above): Add figure showing internal and external relationships
    from rules on action Go. Need a fig from JB}

  Figure xx \kcomment{which figure?} illustrates the entities and
  relations involved in specifying such a rule on a \texttt{Go}
  action. The token on which the rule applies is referred to
  (\texttt{Go}) as the \emph{master}. Each \texttt{At} token required
  by the master is referred to as a \emph{slave}. All variables are
  indicated by name and their domains are expressed as intervals in
  the case of temporal variables and as enumerations for the
  remainder. Application of a rule on a token can thus cause slave
  tokens, variables and constraints to be introduced.  The capability
  of a domain rule to cause a slave token to be created is a key
  vehicle through which planning occurs. Semantically, this rule
  imposes a requirement for supporting (slave) tokens to be in the
  plan in order for the master to be valid.

>>>>

  \item \textit{state}: Tokens can be \texttt{ACTIVE, INACTIVE,
      MERGED}, or \texttt{REJECTED}. The state variable captures the
    token's current state and its reachable states through further
    restriction.  This is required to support \texttt{CAIP}'s approach
    to planning where goals and subgoals can be satisfied by either
    adding new intervals (tokens) to the plan, or by merging with
    existing intervals.

@@

\item[\textbf{Built-in Object Types}] In a domain model, there may be
  some classes that don't have any time-dependent state or behavior,
  and therefore those classes will not have any token types associated
  with them. For instance in the Shopping Agent model, the set
  locations and products are static for the agent's purposes so one
  may only want to declare them but not associate any state or
  behavior with them.  However, a common case in a non-trivial domain
  model is that objects will have many associated tokens in order to
  describe their state and behavior throughout a plan. There are some
  object types that are used so often in domain descriptions that \eu
  provides a built-in implementation for them:

@@

\item[\textbf{Token State Model}] In \eu's representation, goals are
  temporally scoped states that need to be achieved, or actions that
  must be performed. As a result, \eu internally represents goals as
  tokens.  Goals can be posted explicitly, for example, when a user
  asks the Shopping Agent to own a specific product by a specific
  time, or implicitly when subgoals are created as a result of domain
  rules. For instance, when the Shopping Agent must get to a location where a product is
  available before it is able to purchase it: in this case the
  original goal of purchasing the product spawns a subgoal of being at
  a specific location without which the purchase will not occur as
  defined in the domain model.  Looking at how a simple Shopping
  Agent goal would be addresed by \eu helps illustrate how Token
  States support the planning process.  Let's assume the planner
  starts with a partial plan that places the Shopping agent at home at
  time $0$ with no products in possession. Let's also assume that goals are posted for
  the Agent to own a drill by time 10 and to be home by time 20 as depicted in Fig. \ref{fig:europapr1}

\begin{figure} \centering
  \includegraphics[scale=0.45]{figs/europa-pr-1.jpg}
  \caption{\small Timelines representing initial state and goals for a Shopping Agent.}
\label{fig:europapr1}
\vskip+0.1cm
\end{figure}

  Note that the tokens \texttt{Own(Drill)}  and \texttt{At(Home)} that represent the goals are
  not immediately placed on the \texttt{AgentBag} or \texttt{AgentLocation} timelines since \eu
  must first determine whether it can achieve those goal and how. When the
  goals are posted and before any decisions are made, the
  \texttt{Own(Drill)} and \texttt{At(Home)} tokens are created in an \texttt{INACTIVE}
  state. \eu then considers the partial plan and evaluates two
  possible alternatives:

@@

When a token is merged, all restrictions imposed on the merged token
are transferred to the active token upon which it is merged. Merging a
token requires finding a target active token that is compatible. Two
tokens are merge-compatible if they are instances of the same
predicate and there are no empty intersections between corresponding token variables. 
The effect of merging is illustrated in Fig. \ref{fig:europapr:merge}

\begin{figure} \centering
  \includegraphics[scale=0.45]{figs/europa-pr-merge.jpg}
  \caption{\small Timeline states after satisfying goal to be home by time 20 by merging with initial state}
\label{fig:europapr:merge}
\vskip+0.1cm
\end{figure}

The goal to be home by time 20 is satisfied by merging with the initial state that said that the Agent is home at time 0.
The merge is possible because the tokens are of the same type \texttt{At}, and intersecting the domains for all of their variables
doesn't yield an empty domain (which represents an inconsistency), in this case the token variables are the location, 
which is \texttt{Home} in both cases, and the temporal variables (start, end, duration), which can be intersected without causing 
an inconsistency. After the token representing the goal is merged onto the token representing the initial state, the resulting token
represents not only the fact that the Agent must be \texttt{At(Home)} by time 0 but also \texttt{At(Home)} at time 20, this is reflected in 
the new constraints on the token's start and end variables as depicted in Fig. \ref{fig:europaprmerge}.

When a token is activated, all domain rules associated with its type are evaluated,
which may lead to new subgoals being generated. In our example, there
could be a domain rule stating that if a  Shopping Agent doesn't already own a product,
then it must \texttt{Buy} it before it can \texttt{Own} it. In such a case,
activating the \texttt{Own(Drill)} token will cause a new
\texttt{Buy(Drill)} to be created in an \texttt{INACTIVE}
state, which in turn could spawn its own subgoals, like being at the location where the product
can be purchased before being able to buy it, and so on, until all domain rules are satisfied.

\begin{figure} \centering
  \includegraphics[scale=0.45]{figs/europa-pr-activate.jpg}
  \caption{\small Timeline states after satisfying goal to own drill by time 10 by activating corresponding token}
\label{fig:europapr:activate}
\vskip+0.1cm
\end{figure}

It is easy to see in Fig. \ref{fig:europapr:activate} that the decision to merge the goal to be \texttt{At(Home)} by time 20 
with the initial state was not a good one since the Agent will have to get away from \texttt{Home} to be able to 
buy the drill which is inconsistent with the constraints that resulted after the merge. As a result that decision will have 
to be retracted at some point and the alternative to activate or merge with a different token will have to be considered.

If, after looking at all possibilities, neither activating nor merging work, a token may be
\texttt{REJECTED}. This is possible only for goals posted by the end
user and specified as optional. Subgoals generated as a result of
token activation cannot be rejected. If a subgoal cannot be satisfied
through activation or merging then it will cause the original
activation decision that spawned it to be retracted.

As can be seen from the above example, during the search for a plan, a token's state can go from
\texttt{INACTIVE} to \texttt{ACTIVE, MERGED} or \texttt{REJECTED}, as
shown in Fig. \ref{fig:europapr:states}.

\begin{figure} \centering
  \includegraphics[scale=0.45]{figs/europa-pr-states.jpg}
  \caption{\small Possible transitions for a Token State}
\label{fig:europapr:states}
\vskip+0.1cm
\end{figure}

@@

  \begin{figure} \centering
  \includegraphics[scale=0.45]{figs/europa-pr-rules.jpg}
  \caption{\small Master and Slave tokens generated as a result of a Domain Rule}
\label{fig:europapr:rules}
\vskip+0.1cm
\end{figure}

  Fig. \ref{fig:europapr:rules} illustrates the entities and
  relations involved in specifying such a rule on a \texttt{Go}
  action. The token on which the rule applies  (\texttt{Go}) is referred to
 as the \emph{master}. Each \texttt{At} and \texttt{Going} token required
  by the master is referred to as a \emph{slave}. 
  Application of a rule on a token can thus cause slave
  tokens, variables and constraints to be introduced.  The capability
  of a domain rule to cause a slave token to be created is a key
  vehicle through which planning occurs. Semantically, a rule
  imposes a requirement for supporting (slave) tokens to be in the
  plan in order for the master token to be valid.

### 2012-02-22 trunk/europa-basics.tex
added figures missing in concepts.tex

\paragraph \textit{Scheduling}: In the Resource Constrained Project
Scheduling Problem (RCPSP) \cite{Bruckera99}, a project consisting of
a set of activities must be scheduled in a way that satisfies minimum
and/or maximum temporal separation constraints. The activity schedule
must also respect fixed limits on the availability of resources
required to perform each activity. In addition to satisfying temporal
and resource constraints, it is common for the user to want to
minimize makespan so that the entire project is finished as early as
possible. Fig. \ref{fig:rcpsp-1} shows an example of a a solution
provided by \eu for an RCPSP instance with 10 activities, 5 resources
and 30 temporal constraints.

\begin{figure}
\centering
\includegraphics[scale=0.35]{figs/Example-UBO0.jpg}
\caption{\small A \eu solution to an RCPSP \cite{Bruckera99} problem.}
\label{fig:rcpsp-1}
\vskip+0.1cm
\end{figure}

\paragraph \textit{Planning}: In the Shopping Agent Problem
\cite{russelnorvig} an agent needs to purchase a set of products (milk,
drill, etc) that are available at specific locations (supermarket,
hardware store, etc), the agent is subject to temporal (must complete
tasks by specific deadlines) and resource (fuel, carrying capacity, etc)
constraints. The agent needs to figure out what actions need to be
performed to find and acquire the required items, as well as when to
perform each of those actions. Fig. \ref{fig:shopping-1} shows a
solution produced by \eu for such a problem instance.

@@

\item Given that action choices and relationships (rules) are expressed
  through variables, the domain descriptions that result from this
  approach are not intuitive and therefore difficult to understand and
  debug

\item If the structure of a planning model (actions, conditions,
  effects, dependencies at the action level) is not explicitly
  maintained by the CSP planner, the search algorithms are deprived of
  critical information to make better decisions. If that structure is
  maintained (for instance, by internally marking variables that
  represent action choice and relationships between actions), it is
  still hard to write search algorithms as any planning-specific
  insight has to be translated into the CSP representation of
  variables and constraints.

@@

These drawbacks have been addressed by a small number of planning
frameworks like \texttt{Descartes} \cite{joslin95,joslin96},
\texttt{IxTet} \cite{ghallab87,lemai04}, and \texttt{CAIP} (Constraint-Based
Interval and Attribute Planning)\cite{mus94,frank2003}.  These
frameworks take advantage of a CSP representation to deal with time
and resources, at the same time maintaining an explicit representation
of the elements relevant to planning (action choices, causal
relationships, etc). This combination results in a planning approach
that can deal with the type of constraints found in real-life
problems, while allowing for intuitive problem descriptions and a
natural way to express search algorithms and heuristics. \eu is an
implementation of the \texttt{CAIP} framework. In \texttt{CAIP}, planning problems are
specified in terms of:

>>>>

\paragraph \textit{Scheduling}: In the Resource Constrained Project
Scheduling Problem (\texttt{RCPSP}) \cite{Bruckera99}, a project
consisting of a set of activities must be scheduled in a way that
satisfies minimum and/or maximum temporal separation constraints. The
activity schedule must also respect fixed limits on the availability
of resources required to perform each activity. In addition to
satisfying temporal and resource constraints, it is common for the
user to want to minimize makespan \cite{ghallab04} so that the entire
project is finished as early as possible. Fig. \ref{fig:rcpsp-1} shows
an example of a a solution provided by \eu for an \texttt{RCPSP}
instance with 10 activities, 5 resources and 30 temporal constraints.

\begin{figure}
\centering
\includegraphics[scale=0.35]{figs/Example-UBO0.jpg}
\caption{\small A \eu solution to an \texttt{RCPSP} \cite{Bruckera99}
  problem.}
\label{fig:rcpsp-1}
\vskip+0.1cm
\end{figure}

\paragraph \textit{Planning}: The Shopping Agent Problem
\cite{russelnorvig} is a variation of the problem with which we started this chapter.
In this problem, an agent needs to purchase a set of products
(milk, drill, etc) that are available at specific locations
(supermarket, hardware store, etc) and the agent is subject to
temporal (must complete tasks by specific deadlines) and resource
(fuel, carrying capacity, etc) constraints. The agent needs to figure
out what actions need to be performed to find and acquire the required
items, as well as when to perform each of those
actions. Fig. \ref{fig:shopping-1} shows a solution produced by \eu
for such a problem instance.

@@

\item Given that action choices and relationships (rules) are
  expressed through variables, the domain descriptions that result
  from this approach are not intuitive and therefore difficult to
  understand and debug.

\item If the structure of a planning model (actions, conditions,
  effects, dependencies at the action level) is not explicitly
  maintained by the \textsf{CP} planner, the search algorithms are deprived of
  critical information to make better decisions. If that structure is
  maintained (for instance, by internally marking variables that
  represent action choice and relationships between actions), it is
  still hard to write search algorithms as any planning-specific
  insight has to be translated into the \textsf{CP} representation of
  variables and constraints.

@@

These drawbacks have been addressed by a number of planning frameworks
like \texttt{Descartes} \cite{joslin95,joslin96}, \texttt{IxTet}
\cite{ghallab87,lemai04} and \texttt{CAIP} (Constraint-Based Interval
and Attribute Planning) \cite{mus94,frank2003}.  These frameworks take
advantage of a \textsf{CP} representation to deal with time and resources, at
the same time maintaining an explicit representation of the elements
relevant to planning (action choices, causal relationships, etc). This
combination results in a planning approach that can deal with the type
of constraints found in real-world problems while allowing for
intuitive problem descriptions and a natural way to express search
algorithms and heuristics. \eu is one such implementation of the
\texttt{CAIP} framework, in which planning problems are specified in
terms of:
